---
title: "Getting data about trains in India"
author: "Maëlle Salmon"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---
Train stations and train in India, data issues
================================================

# Introduction

For the CHAI project <http://www.chaiproject.org/> I wanted to look whether there were trains in our study area. Besides, I was curious about two data sources:

* The Open Government Data (OGD) Platform of India <https://data.gov.in/>

* Openstreetmap <https://www.openstreetmap.org/>

I moreover saw this blog article <http://curiousanalytics.blogspot.com.es/2015/04/indian-railways-network-using-r.html> 

In this README I'll explain where I got the data I decided to use, and which problems I still have.

# Getting and preparing the timetable

I used a different source than curiousanalytics. I used this train timetable from the OGD Platform of India: <https://data.gov.in/catalog/indian-railways-train-time-table-0> I tried to download the data from the API using the `ogdindiar` package <https://github.com/steadyfish/ogdindiar> but I got an error message from the server so I downloaded the data by hand. I then prepared it using the code in [this R code](R_code/getting_train_timetable.R)

The data look like this:

```{r, echo=TRUE, warning = FALSE, message = FALSE}
load("train_data/train_timetable.RData")
knitr::kable(head(timetable))
```

It has `r nrow(timetable)` rows.

Now, how to get geographical coordinates for each station?

# Getting the coordinates for all stations

## Google Maps info 

I first used the same approach as curiousanalyics: querying Google Maps via the `geocode` function of the `ggmap` package for getting coordinates for each station name. One can see the code used for this at the beginning of [this R code](R_code/getting_coordinates.R)

After doing this I only had geographical coordinates for about 40% of the train stations.

## Openstreetmap info

Then I decided to get all train stations nodes from Openstreetmap. I downloaded the osm.pbf file for India from <http://download.geofabrik.de/asia/india.html> I filtered only the "railway=station" from this file using osmosis. The script is [here](osm_data/osmosis_script.txt). The osm.pbf file is not there because it was too big.

I parsed the OSM XML file using [this code](osm_data/reading_osm_file.R). Here maybe I could have made better use of the xml2 package and also of the osmar package but somehow I found it faster to write this code.

The data look like this:

```{r, echo=TRUE, warning = FALSE, message = FALSE}
load("osm_data/OSMdataIndiaStations.RData")
knitr::kable(head(dataIndiaStations))
```

It has `r nrow(dataIndiaStations)` rows. So many train stations!

I gave priority to info I got from Google Maps but for the remaining ones I looked for close matches for names in the OSM data. I defined closes matches as names whose difference measured by the `stringdist` function of the `stringdist` package was 0 or 1. It is a bad solution because clearly doing this I'm giving wrong coordinates to some stations. 

The whole code is [here](R_code/getting_coordinates.R)

The resulting data look like this:

```{r, echo=TRUE, warning = FALSE, message = FALSE}
load("geo_data/geoInfo.RData")
knitr::kable(head(listNames))
```

It has `r nrow(listNames)` rows.

# Remaining issues

* How to get coordinates for all stations?

* In general, how to optimally deal with different spellings of Indian locations?

* How should I read OSM files in a more elegant way? 

# Cool stuff I've learnt

* Openstreetmap and the OGD platform of India are goldmines.

* Comparing strings of characters could help with the spelling of Indian locations but also generally in our questionnaire data for dealing with typing errors in the free text areas.

# Acknowledgements

I've read many forums in order to understand e.g. Openstreetmap, so thank you to all the people that asked and answered questions on these forums! Here is a smiling cat for all of you nice people. :smile_cat: